---
title: "Practical Machine Learning Project"
author: "Anh"
date: "2/4/2022"
output: pdf_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Background
Using devices such as Jawbone Up, Nike FuelBand, and Fitbit is now possible to 
collect a large amount of data about personal activity relatively inexpensively. 
These type of devices are part of the quantified self movement – a group of 
enthusiasts who take measurements about themselves regularly to improve their 
health, to find patterns in their behavior, or because they are tech geeks. One 
thing that people regularly do is quantify how much of a particular activity 
they do, but they rarely quantify how well they do it.

# Data processing
## Importing data
```{r message = FALSE, warning = FALSE}
# Load necessary packages
library(caret)
library(rattle)
library(rpart)
library(rpart.plot)
library(randomForest)
library(repmis)
```
```{r message = FALSE, warning = FALSE}
# Load data locally
df_train = read.csv("pml-training.csv", na.strings = c("NA", ""))
df_test = read.csv("pml-testing.csv", na.strings = c("NA", ""))
dim(df_train)
dim(df_test)
```
From the dim functions, we can see that the training dataset has 19622 
observations and 160 variables, and the testing data set contains 20 
observations and the same variables as the training set. We are trying to 
predict the outcome of the variable `classe` in the training set.

## Cleaning data
```{r message = FALSE, warning = FALSE}
# delete columns of the training set that contain any missing values
train = df_train[, colSums(is.na(df_train)) == 0]
test = df_test[, colSums(is.na(df_test)) == 0]

# delete the first seven weak predictors 
train = train[, -c(1:7)]
test = test[, -c(1:7)]
```
## Spliting data
To get out-of-sample errors, we split the cleaned training set `train` into a 
training set (`train_data`, 70%) for prediction and a validation set 
(`valid_data` 30%) to compute the out-of-sample errors.

```{r message = FALSE, warning = FALSE}
set.seed(7826) 
inTrain <- createDataPartition(train$classe, p = 0.7, list = FALSE)
train_data <- train[inTrain, ]
valid_data <- train[-inTrain, ]
```
# Prediction algorithms
## Classification trees
Since data transformations are not necessarily important in non-linear models 
like classification trees, we do not transform any variables.

```{r message = FALSE, warning = FALSE}
control <- trainControl(method = "cv", number = 5)
fit_rpart <- train(classe ~ ., data = train_data, method = "rpart", 
                   trControl = control)
print(fit_rpart, digits = 4)
```
```{r message = FALSE, warning = FALSE}
fancyRpartPlot(fit_rpart$finalModel)
```
```{r message = FALSE, warning = FALSE}
# predict outcomes using validation set
predict_rpart <- predict(fit_rpart, valid_data)
# Show prediction result
(conf_rpart <- confusionMatrix(as.factor(valid_data$classe), predict_rpart))
```

```{r message = FALSE, warning = FALSE}
(accuracy_rpart <- conf_rpart$overall[1])
```
Accuracy rate is 0.5, and so the out-of-sample error rate is 0.5. Therefore, 
using classification tree does not predict the outcome `classe` very well.

## Random forests

```{r message = FALSE, warning = FALSE}
fit_rf <- train(classe ~ ., data = train_data, method = "rf", 
                   trControl = control)
print(fit_rf, digits = 4)
```

```{r message = FALSE, warning = FALSE}
# predict outcomes using validation set
predict_rf <- predict(fit_rf, valid_data)
# Show prediction result
(conf_rf <- confusionMatrix(as.factor(valid_data$classe), predict_rf))
```

```{r message = FALSE, warning = FALSE}
(accuracy_rf <- conf_rf$overall[1])
```
The accuracy rate is 0.991, and so the out-of-sample error rate is 0.009. This 
may be due to the fact that many predictors in the dataset are highly correlated. 
Random forests choose a subset of predictors at each split and decorrelate the 
trees. This leads to high accuracy, although this algorithm is sometimes 
difficult to interpret and computationally inefficient.

# Prediction on test set
```{r message = FALSE, warning = FALSE}
(predict(fit_rf, test))
```





